Transient Corotating Clumps Around Adolescent Low-Mass Stars From Four Years of TESS

### 2023-09-07 paper/ms.tex
update ms

% 250 word max!
\begin{abstract}
  %
  Complex periodic variables (CPVs) are stars that exhibit sharp and
  periodic optical light curves.  Previous studies have indicated that
  these stars are typically disk-free pre-main-sequence M dwarfs with
  rotation periods ranging from 0.2 to 2 days.  To advance our
  understanding of these enigmatic objects, we conducted a blind
  search using TESS 2-minute data of \nstarssearched\ K and M dwarfs
  with $T$$<$16 and $d$$<$150\,pc.  We found \ngoods\ high-quality
  CPVs, and subsequently determined that most are members
  of stellar associations.  Among the new discoveries are the
  brightest ($T$$\approx$9.5), closest ($d$$\approx$20\,pc), and
  oldest ($\approx$200\,Myr) CPVs known.  One exceptional object, LP
  12-502, exhibited up to eight flux dips per cycle. Some of these
  dips coexisted with slightly different periods, and the
  shortest-duration LP 12-502 dips precisely matched the expected
  timescale for transiting small bodies at the corotation radius.
  Broadly, our search confirms that CPVs are mostly young
  ($\lesssim$150\,Myr) and low-mass ($\lesssim$0.4\,$M_\odot$).  The
  flux dips characteristic of the class have lifetimes of $\approx$100
  cycles, although stellar flares seem to induce sudden dip collapse
  once every few months.  The most plausible explanation for these
  phenomena remains corotating concentrations of gas or dust.  The gas
  or dust is probably entrained by the star's magnetic field, and the
  sharp features could result from a multipolar field topology, a
  hypothesis supported by correspondences between the light curves of
  CPVs and of rapidly-rotating B stars known to have multipolar magnetic fields.
  %
\end{abstract}

@@

{\sc Dip durations}---The shortest dip duration for any of the
individual LP 12-502 dips seems to be $\approx$0.06\,$P$ $\approx$
1.08\,hr.  This is very similar to the characteristic timescale of a
transiting small body at the corotation radius,
\begin{equation}
T_{\rm dur} \equiv R_\star P_{\rm rot} / (\pi a) = 1.02\pm 0.10\,{\rm hr},
\end{equation}
where we have inserted the stellar radius and mass derived in
Section~\ref{subsec:starprops}.  Thus, the shortest-duration dips
could be produced by transits of bodies or distributions of material
that are smaller than the star.  The corotation radius corresponds to
$a/R_\star \approx 5.8$, i.e., the transit of a body at the corotation
radius has a duration about six times shorter than a feature on the
stellar photosphere that is carried across the visible hemisphere by
rotation.  On the other hand, some dip durations are sufficiently long
that an explanation involving transits would require structures that
are larger than the star along the direction of orbital motion.

@@

\subsection{Strengthening the magnetic B star connection}

@@

The physical similarity between the B stars and the M dwarfs could
have its origin in the existence of a ``centrifugal magnetosphere''
\citep[see][]{2013MNRAS.429..398P}.  In other words, both classes of
objects might satisfy the condition $R_{\rm m} > R_{\rm c}$, for
$R_{\rm m}$ the magnetosphere radius (sometimes called the Alfv\'en
radius).  Provided that charged particles are confined to move along
magnetic field lines, material can then build up at the corotation
radius \citep[see][Sec.~4, and references
therein]{2015SSRv..191..339R}.  In the converse ``dynamical'' case,
when $R_{\rm m} < R_{\rm c}$, material interior to the magnetospheric
radius returns to the stellar surface over the free-fall timescale.  A
simple estimate assuming a dipole field with $B_0\approx1\,{\rm kG}$
at the star's surface, a local plasma number density
$n\approx10^9\,{\rm cm^{-3}}$, and a plasma temperature $10^6\,{\rm
K}$ gives magnetospheric radii of order a few times the corotation
radii, $R_{\rm c}$.  This suggests that the existence of a centrifugal
magnetosphere is plausible for young, rapidly rotating M dwarfs.

@@

We found \ngoods\ objects that showed complex
quasiperiodic behavior over at least one TESS sector.  These \ngoods\
bona fide CPVs are listed in Table~\ref{tab:thetable}.  This table also
includes \nmaybes\ ambiguous CPVs, whose designation is
less certain, and \ndebunked\ impostors.  We inferred ages for all but
two of the \nallcands\ objects based on memberships in young stellar assocations; we
also derived temperatures and radii using SED fitting, and 
inferred stellar masses by interpolating against stellar evolutionary
models.  We caution that our sample is far from being
volume-limited and is not even magnitude-limited:
the TESS 2-minute stellar sample had a heterogeneous
selection function which may have been biased in favor of young stars
over field stars.  Previous work however has shown that $\approx$1-3\%
of M dwarfs younger than $\approx$100\,Myr show the CPV phenomenon
\citep{2016AJ....152..114R,2022AJ....163..144G,2022AJ....164...80R}.

@@

      Such dips may therefore be produced by transits of bodies or

@@

      inhomogeneities in the stellar surface brightness
      distribution (starspots and faculae).  The

@@

      correspond to dips collapsing instantaneously, and occur once
      every few months for both LP~12-502 and TIC~300651846.  They are
      almost always linked with observed optical flares.  Such
      switches are suggestive of magnetic reconnection opening the
      ``magnetic cage'' that traps the dust.

@@

    \item The on-off duty cycle for CPVs seems to be $\approx$$75$\%, based

@@

      moving group (TIC~294328887; $\approx$200\,Myr).  However the

@@

      limit.  However both are potentially binaries, and this may

@@

    \item The closest CPVs to the Sun are at distances of 15-20\,pc,
      and the brightest have $V$$\approx$12 ($J$$\approx$7.5).  We
      have found most of them in this work, since our CPV sample was

@@

    \item Surprising analogs to the CPVs exist in two magnetic B stars,

@@

      topology.  Since most magnetic B stars have dipolar magnetic
      fields \citep{2009ARA&A..47..333D}, this suggests that the CPV

>>>>

% 250 word max!
\begin{abstract}
  %
  Complex periodic variables (CPVs) are stars that exhibit sharp and
  periodic optical light curves.  Previous studies have indicated that
  these stars are typically disk-free pre-main-sequence M dwarfs with
  rotation periods ranging from 0.2 to 2 days.  To advance our
  understanding of these enigmatic objects, we conducted a blind
  search using TESS 2-minute data of \nstarssearched\ K and M dwarfs
  with $T$$<$16 and $d$$<$150\,pc.  We found \ngoods\ high-quality
  CPVs, and subsequently determined that most are members
  of stellar associations.  Among the new discoveries are the
  brightest ($T$$\approx$9.5), closest ($d$$\approx$20\,pc), and
  oldest ($\approx$200\,Myr) CPVs known.  One exceptional object, LP
  12-502, exhibited up to eight flux dips per cycle. Some of these
  dips coexisted with slightly different periods, and the
  shortest-duration LP 12-502 dips precisely matched the expected
  timescale for transiting small bodies at the corotation radius.
  Broadly, our search confirms that CPVs are mostly young
  ($\lesssim$150\,Myr) and low-mass ($\lesssim$0.4\,$M_\odot$).  The
  flux dips characteristic of the class have lifetimes of $\approx$100
  cycles, although stellar flares seem to induce sudden dip collapse
  once every few months.  The most plausible explanation for these
  phenomena remains corotating concentrations of gas or dust.  The gas
  or dust is probably entrained by the star's magnetic field, and the
  sharp features could result from a multipolar field topology, a
  hypothesis supported by correspondences between the light curves of
  CPVs and of rapidly rotating B stars known to have multipolar magnetic fields.
  %
\end{abstract}

@@

{\sc Dip durations}---The shortest dip duration for any of the
individual LP 12-502 dips seems to be $\approx$0.06\,$P$ $\approx$
1.08\,hr.  This is very similar to the characteristic timescale of a
transiting small body at the corotation radius,
\begin{equation}
T_{\rm dur} \equiv R_\star P_{\rm rot} / (\pi a) = 1.02\pm 0.10\,{\rm hr},
\end{equation}
where we have inserted the stellar radius and mass derived in
Section~\ref{subsec:starprops}.  Thus, the shortest-duration dips
are likely produced by transits of bodies or distributions of material
that are smaller than the star.  The corotation radius corresponds to
$a/R_\star \approx 5.8$, i.e., the transit of a body at the corotation
radius has a duration about six times shorter than a feature on the
stellar photosphere that is carried across the visible hemisphere by
rotation.  On the other hand, some dip durations are sufficiently long
that an explanation involving transits would require structures that
are larger than the star along the direction of orbital motion.

@@

\subsection{Strengthening the magnetic B star connection}
\label{subsec:bstardisc}

@@

The physical similarity between the B stars and the M dwarfs could
have its origin in the existence of a ``centrifugal magnetosphere''
\citep[see][]{2013MNRAS.429..398P}.  In other words, both classes of
objects might satisfy the condition $R_{\rm m} > R_{\rm c}$, for
$R_{\rm m}$ the magnetosphere radius (sometimes called the Alfv\'en
radius).  Provided that charged particles are confined to move along
magnetic field lines, material can then build up at the corotation
radius \citep[e.g.][Sec.\ 4, and references
therein]{2015SSRv..191..339R}.  In the converse ``dynamical'' case,
when $R_{\rm m} < R_{\rm c}$, material interior to the magnetospheric
radius returns to the stellar surface over the free-fall timescale.  A
simple estimate assuming a dipole field with $B_0\approx1\,{\rm kG}$
at the star's surface, a local plasma number density
$n\approx10^9\,{\rm cm^{-3}}$, and a plasma temperature $10^6\,{\rm
K}$ gives magnetospheric radii of order a few times the corotation
radii, $R_{\rm c}$.  This suggests that the existence of a centrifugal
magnetosphere is plausible for young, rapidly rotating M dwarfs.

@@

We found \ngoods\ objects that showed complex
quasiperiodic behavior over at least one TESS sector.  These \ngoods\
bona fide CPVs are listed in Table~\ref{tab:thetable}.  This table also
includes \nmaybes\ ambiguous CPVs, whose designation is
less certain, and \ndebunked\ impostors.  We inferred ages for all but
two of the \nallcands\ objects based on memberships in young stellar associations; we
also derived temperatures and radii using SED fitting, and 
inferred stellar masses by interpolating against stellar evolutionary
models.  We caution that our sample is far from being
volume-limited and is not even magnitude-limited:
the TESS 2-minute stellar sample had a heterogeneous
selection function which may have been biased in favor of young stars
over field stars.  Previous work however has shown that $\approx$1-3\%
of M dwarfs younger than $\approx$100\,Myr show the CPV phenomenon
\citep{2016AJ....152..114R,2022AJ....163..144G,2022AJ....164...80R}.

@@

      Such dips are therefore likely produced by transits of bodies or

@@

      starspots and faculae on the stellar surface.  The

@@

      correspond to dips collapsing instantaneously;  they occur once
      every few months for both LP~12-502 and TIC~300651846.
      State-switches are almost always linked with observed optical
      flares.  Such switches are suggestive of magnetic reconnection
      opening the ``magnetic cage'' that traps the dust.

@@

    \item The on-off duty cycle for CPVs is $\approx$$75$\%, based

@@

      moving group (TIC~294328887; $\approx$200\,Myr).  The

@@

      limit.  Both are potentially binaries, and this may

@@

    \item The closest CPVs to the Sun are at distances of 15-20\,pc;
      the brightest have $V$$\approx$12 ($J$$\approx$7.5).  We
      have found most of the close exemplars in this work, since our CPV sample was

@@

    \item Surprising analogs to CPVs exist in two magnetic B stars,

@@

      topology (Section~\ref{subsec:bstardisc}).  Since most magnetic
      B stars have dipolar magnetic fields, this suggests that the CPV
